Hybrid Partial Least Squares Regression with Multiple Functional and Scalar Predictors

### 2025-08-21 manuscript/main.tex
adjoint

	\subsection{ Tucker's Criterion  }\label{section:sub:tucker} 
	    \jongmin{comment:maybe we can move this section to precede the algorithm section. then the new structure will introduce population version first and introduce finite-sample estimation procedure.}
	The fundamental property of classical scalar PLS is that the PLS component direction  corresponds to the first eigenvector of the squared covariance matrix, satisfying  the Tucker's Criterion \citep{Tucker1938}.
	Under a mild assumption,
	We derive an analogous result  in our scalar-on-hybrid regression model setting \eqref{eq: Hybrid functional model}.
	This justifies our estimation procedure introduced in Section \ref{section:main:our_algorithm}.
For notational simplicity, we omit the index $l$. At the population level, where the random objects are fully observed, it suffices to consider $Y_1$ and $W_1$ rather than their sampled realizations.
We assume 
$\mathbb{E}[Y_1]=0 \in \mathbb{R}$ and $\mathbb{E}[W_1] = 0 \in \mathbb{H}$.
 We first define the cross-covariance between the response and the functional predictors, where the predictors correspond to the observed versions for $l=1$ and to the residualized versions for $l \ge 2$:
	\begin{equation}\label{def:cross_covariance_yx}
		\sigma_{YX}
		=
		\bigl(
		\sigma_{YX, 1}(t) , \ldots, \sigma_{YX, K}(t) 
		\bigr)
		:=
		\bigl(
		\mathbb{E}
		[Y_1 X_{11}]
		,\ldots, 
		\mathbb{E}
		[Y_1  X_{1K}]
		\bigr)   \in \mathcal{F}^{\otimes K},
	\end{equation}
	and between the response and the scalar predictor:
	\begin{equation*}
		\sigma_{YZ} 
		= 
		(\sigma_{YZ,1}, \ldots, \sigma_{YZ,p})^\top 
		:= 
		\bigl(
		\mathbb{E}[Y_1 Z_{11}]
		,\ldots,
		\mathbb{E}[Y_1 Z_{1p}]
		\bigr)^\top %= E(Y\mathbf{Z})
		\in \mathbb{R}^p.
	\end{equation*}
	Based on these definitions, we define the cross-covariance term between the response and the hybrid predictor as
	\begin{equation*}
		\Sigma_{YW} :=  
		\mathbb{E}
		[ Y_1 W_1 ]
		=
		\bigl(
		\sigma_{YX, 1} , \ldots, \sigma_{YX, K} , \, \sigma_{YZ}
		\bigr) \in \mathbb{H}.
	\end{equation*}
	Based on these definitions, we introduce two cross-covariance operators and their properties.
	\begin{lemma}\label{lemma:cross_cov_functional}
		Define the operator $
		\mathcal{C}_{YW}
		=
		\mathbb{E}(W_1 \otimes_\mathbb{H} Y_1)
		: \mathbb{H} \rightarrow \mathbb{R}$ such that for any $h=(f_1, \ldots, f_K, \mathbf{v}) \in \mathbb{H}$, it maps $h$ to a real number as: \begin{equation*} \mathcal{C}_{YW} h := \mathbb{E} \left[ \langle W_1, h \rangle_\mathbb{H} Y_1 \right] = \langle \Sigma_{YW}, h \rangle_\mathbb{H} = \sum \limits_{k=1}^K \int_0^1 \sigma_{YX, k}(t) \, f_k(t) \, dt + \sigma_{YW}^\top \mathbf{v}. \end{equation*} 
		\jongmin{instead of showing $\mathcal{U}$ is compact, let's show $\mathcal{C}_{YW}$ is compact. This way, we don't need uniform continuiity of $\sigma_{YX}$.}If there exist finite constants $Q_1$ and $Q_2$ such that
		\begin{align}\label{condition_for_compact}
			\max \limits_{k=1, \ldots, K} \sup \limits_{t \in [0,1]} \sigma_{YX, k}^{2}(t) < Q_1 \quad \textnormal{and} \quad  \max_{r=1, \ldots, p} \sigma_{YZ, r}^2 < Q_2,
		\end{align}
		the operator $\mathcal{C}_{YW}$ is a compact operator.
	\end{lemma}
	Proof of Lemma \ref{lemma:cross_cov_functional}
	is provided in Appendix \ref{section:proof:lemma:cross_cov_functional}.
	\begin{lemma}\label{lemma:cross_cov_operator}
		Define $\mathcal{C}_{WY} = \mathbb{E}[Y \otimes W]: \mathbb{R} \rightarrow \mathbb{H}$,
		which	maps  $d \in \mathbb{R}$ to a hybrid object in $\mathbb{H}$ as follows:
		%%%%
		\begin{align*}		
			\mathcal{C}_{WY} d 
			:= 
			\mathbb{E} 
			\bigl[
			\langle Y_1, d \rangle W_1 
			\bigr] 
			= 
			\mathbb{E}
			[ Y_1 W_1 d ] =  d \, \Sigma_{Y W}.
		\end{align*}
		Then	$\mathcal{C}_{WY}$ is an adjoint operator of $\mathcal{C}_{YW}$. That is, $\mathcal{C}_{WY} = \mathcal{C}_{YW}^\ast$.
	\end{lemma}

@@

 We define an operator $\mathcal{T}$ that maps an element $g \in \mathbb{H}$   (representing a potential regression coefficient) to the predicted response.
\begin{lemma}\label{lemma:tau_ast}
	Define the linear operator $\mathcal{T}: \mathbb{H} \to \mathbb{R}$ that maps  $g \in \mathbb{H}$ to 
	$
	\mathcal{T}g = \langle W, g \rangle_{\mathbb{H}}
	$.
 Let $\mathcal{T}^\ast: \mathbb{R} \to \mathbb{H}$ denote its adjoint operator. Then,
	\[
	\Sigma_{YW}^{[1]} = \mathcal{T}^\ast Y.
	\]
\end{lemma}
Proof of Lemma \ref{lemma:tau_ast} is provided in Appendix \ref{section:rpoof:lemma:tau_ast}.

@@

\subsection{Proof of Lemma \ref{lemma:tau_ast} }\label{section:rpoof:lemma:tau_ast}
	\begin{proof}
	The adjoint operator  is uniquely defined by the property relating the inner products in $L^2(\Omega)$ and $\mathbb{H}$: $$ \langle \mathcal{T}g, Y \rangle_{L^2(\Omega)} = \langle g, \mathcal{T}^\ast Y \rangle_{\mathbb{H}}, \quad \text{for all } g \in \mathbb{H}, Y \in L^2(\Omega) $$ Here, $\langle \cdot, \cdot \rangle_{L^2(\Omega)}$ denotes the inner product in $L^2(\Omega)$, which is the expectation $E[\cdot]$. Let's explicitly derive the expression for $\mathcal{T}^\ast Y$: $$ \langle g, \mathcal{T}^\ast Y \rangle_{\mathbb{H}} = E[\langle W, g \rangle_{\mathbb{H}} Y] $$ By the properties of expectation and the linearity of the inner product (assuming appropriate boundedness and measurability for interchange of expectation and inner product): $$ E[\langle W, g \rangle_{\mathbb{H}} Y] = E[\langle g, W \rangle_{\mathbb{H}} Y] = \langle g, E[YW] \rangle_{\mathbb{H}} $$ Comparing the leftmost and rightmost parts of the definition of the adjoint: $$ \mathcal{T}^\ast Y = E[YW] $$
	\end{proof}

>>>>

	\subsection{ Tucker's Criterion  }\label{section:sub:tucker} 
	    \jongmin{comment:maybe we can move this section to precede the algorithm section. then the new structure will introduce population version first and introduce finite-sample estimation procedure.}
	The fundamental property of classical scalar PLS is that the PLS component direction  corresponds to the first eigenvector of the squared covariance matrix, satisfying  the Tucker's Criterion \citep{Tucker1938}.
	Under a mild assumption,
	We derive an analogous result  in our scalar-on-hybrid regression model setting \eqref{eq: Hybrid functional model}.
	This justifies our estimation procedure introduced in Section \ref{section:main:our_algorithm}.
For notational simplicity, we omit the index $l$. At the population level, where the random objects are fully observed, it suffices to consider $Y_1$ and $W_1$ rather than their sampled realizations.
We assume 
$\mathbb{E}[Y_1]=0 \in \mathbb{R}$ and $\mathbb{E}[W_1] = 0 \in \mathbb{H}$.
 We first define the cross-covariance between the response and the functional predictors, where the predictors correspond to the observed versions for $l=1$ and to the residualized versions for $l \ge 2$:
	\begin{equation}\label{def:cross_covariance_yx}
		\sigma_{YX}
		=
		\bigl(
		\sigma_{YX, 1}(t) , \ldots, \sigma_{YX, K}(t) 
		\bigr)
		:=
		\bigl(
		\mathbb{E}
		[Y_1 X_{11}]
		,\ldots, 
		\mathbb{E}
		[Y_1  X_{1K}]
		\bigr)   \in \mathcal{F}^{\otimes K},
	\end{equation}
	and between the response and the scalar predictor:
	\begin{equation*}
		\sigma_{YZ} 
		= 
		(\sigma_{YZ,1}, \ldots, \sigma_{YZ,p})^\top 
		:= 
		\bigl(
		\mathbb{E}[Y_1 Z_{11}]
		,\ldots,
		\mathbb{E}[Y_1 Z_{1p}]
		\bigr)^\top %= E(Y\mathbf{Z})
		\in \mathbb{R}^p.
	\end{equation*}
	Based on these definitions, 
	Based on these definitions, we introduce two cross-covariance operators and their properties.

 We define an operator $\mathcal{T}$ that maps an element $g \in \mathbb{H}$   (representing a potential regression coefficient) to the predicted response.
\begin{lemma}\label{lemma:tau_ast}
	Let the cross-covariance between the response $Y$ and the hybrid predictor $W$ be
	\[
	\Sigma_{YW} := \mathbb{E}[ Y_1 W_1 ]
	= \bigl( \sigma_{YX,1}, \ldots, \sigma_{YX,K}, \, \sigma_{YZ} \bigr) 
	\in \mathbb{H}.
	\]
	Define the linear operator $\mathcal{T}: \mathbb{H} \to \mathbb{R}$ by
	\[
	\mathcal{T} g := \langle W_1, g \rangle_{\mathbb{H}}, \quad g \in \mathbb{H}.
	\]
	Let $\mathcal{T}^\ast: \mathbb{R} \to \mathbb{H}$ denote its adjoint operator. Then,
	\[
	\mathbb{E}[ \mathcal{T}^\ast Y_1 ] = \Sigma_{YW}.
	\]
\end{lemma}

Proof of Lemma \ref{lemma:tau_ast} is provided in Appendix \ref{section:rpoof:lemma:tau_ast}.

	\begin{lemma}\label{lemma:cross_cov_functional}
		Define the operator $
		\mathcal{C}_{YW}
		=
		\mathbb{E}(W_1 \otimes_\mathbb{H} Y_1)
		: \mathbb{H} \rightarrow \mathbb{R}$ such that for any $h=(f_1, \ldots, f_K, \mathbf{v}) \in \mathbb{H}$, it maps $h$ to a real number as: \begin{equation*} \mathcal{C}_{YW} h := \mathbb{E} \left[ \langle W_1, h \rangle_\mathbb{H} Y_1 \right] = \langle \Sigma_{YW}, h \rangle_\mathbb{H} = \sum \limits_{k=1}^K \int_0^1 \sigma_{YX, k}(t) \, f_k(t) \, dt + \sigma_{YW}^\top \mathbf{v}. \end{equation*} 
		\jongmin{instead of showing $\mathcal{U}$ is compact, let's show $\mathcal{C}_{YW}$ is compact. This way, we don't need uniform continuiity of $\sigma_{YX}$.}If there exist finite constants $Q_1$ and $Q_2$ such that
		\begin{align}\label{condition_for_compact}
			\max \limits_{k=1, \ldots, K} \sup \limits_{t \in [0,1]} \sigma_{YX, k}^{2}(t) < Q_1 \quad \textnormal{and} \quad  \max_{r=1, \ldots, p} \sigma_{YZ, r}^2 < Q_2,
		\end{align}
		the operator $\mathcal{C}_{YW}$ is a compact operator.
	\end{lemma}
	Proof of Lemma \ref{lemma:cross_cov_functional}
	is provided in Appendix \ref{section:proof:lemma:cross_cov_functional}.
	\begin{lemma}\label{lemma:cross_cov_operator}
		Define $\mathcal{C}_{WY} = \mathbb{E}[Y \otimes W]: \mathbb{R} \rightarrow \mathbb{H}$,
		which	maps  $d \in \mathbb{R}$ to a hybrid object in $\mathbb{H}$ as follows:
		%%%%
		\begin{align*}		
			\mathcal{C}_{WY} d 
			:= 
			\mathbb{E} 
			\bigl[
			\langle Y_1, d \rangle W_1 
			\bigr] 
			= 
			\mathbb{E}
			[ Y_1 W_1 d ] =  d \, \Sigma_{Y W}.
		\end{align*}
		Then	$\mathcal{C}_{WY}$ is an adjoint operator of $\mathcal{C}_{YW}$. That is, $\mathcal{C}_{WY} = \mathcal{C}_{YW}^\ast$.
	\end{lemma}

@@

@@

\subsection{Proof of Lemma \ref{lemma:tau_ast} }\label{section:rpoof:lemma:tau_ast}
	\begin{proof}
By the definition of the adjoint operator, 
for all $g \in \mathbb{H}$, we have
\begin{align*}
	 \langle
	\Sigma_{YW}, g \rangle_{\mathbb{H}} 
=	
	 \langle
	 \mathbb{E}
	 [ Y_1 W_1], g \rangle_{\mathbb{H}} 
	=
\mathbb{E}
[ Y_1 \langle W_1, g \rangle_{\mathbb{H}} ]
=
\mathbb{E}[
Y_1
(\mathcal{T}g)
]
=
\mathbb{E}[
\langle \mathcal{T}g, Y_1 \rangle
]
&= \mathbb{E}[
\langle g, \mathcal{T}^\ast Y_1 \rangle_{\mathbb{H}}
]
\\&= 
\langle g, \mathbb{E}[\mathcal{T}^\ast Y_1 ]\rangle_{\mathbb{H}}.
\end{align*}	
This completes the proof of Lemma \ref{lemma:tau_ast}.
	\end{proof}
